Nonfungible tokens allow \href{https://en.wikipedia.org/wiki/List_of_most_expensive_non-fungible_tokens}{digital and new media artists} to connect with audiences without gatekeepers. Established mediators and curators of art have been caught totally wrongfooted, and NFTs seem to give a way for them to be cut out completely. There are suggestions of applications beyond this initial digital art scope. This is a compounding, and disrupting paradigm change.\par
%Users of NFT markets have \href{https://blog.chainalysis.com/reports/nft-market-report-preview-2021/}{injected around \$30 billion into the tokens during 2021}. 
While it is likely that there is currently a speculative bubble, and indeed is waning already, it seems certain that the technology is here to stay.\par
%Samsung for instance have announced that their TVs will support not only \href{https://news.samsung.com/us/samsung-2022-micro-led-neo-qled-lifestyle-tvs-personalization-options-ces-2022/}{display of NFTs} with artist defined settings in the metadata, but also an integrated marketplace for browsing and purchasing.\par
One of the most interesting companies is Yuga Labs, who launched the incredibly popular Bored Ape Yacht club set of 10,000 algorithmically generated NFTs. These Ethereum based NFTs regularly change hands for hundreds of thousands of pounds.\par
Yuga recently bought the artistic rights to the commercial reuse of similarly popular `Crypto Punks` set. This is interesting because they have handed the commercial re-use rights to the owners of the individual NFTs. \par
The community around these collections is incredibly strong, mixing developers, artists, the rich and famous, and the fortunate and early, into a cohesive community who communicate online. The developer `good will' is enormous, and it seems possible that this will lead to faster and broader innovation around the collections, and out into \href{https://twitter.com/yugalabs/status/1505014986556551172?}{metaverse applications}. The brand is strong, and the individual NFT items both benefit from, and reinforce that brand, while adding personal narratives and human interest.\par 
As a gauge of how frothy this market still is it's interesting to look at the APE token which Yuga just launched. They airdropped 10,000 of the tokens free to each of the 10,000 NFT holders. This instantly created a multi-billion dollar market cap, and a top 50 `crypt' out of thin air. It's clear that there is both brand, and a market here.\par



%\href{https://amycastor.com/2021/03/14/metakovan-the-mystery-beeple-art-buyer-and-his-nft-defi-scheme/}{Beeple scam}

\section{Digital Art}
\input{07_1_NFTUmran}
%\section{NFTs and games}

\section{Other uses}
Peter Thiel, the billionaire venture capitalist who founded PayPal has invested in expanded NTF use cases. The first is `Royal' which is experimentally  selling \href{https://royal.io/}{selling limited NFT tokens} which contractually entitle the holder to a portion of music artist royalties. The other is a \href{https://www.ztonft.com/}{political funding NFT} from Blake Masters to support his senate ambitions.

It is completely reasonable to assert that these use cases could be accomplished without the use of NFT technology, and is part of the hype bubble.\par
NFT art currently suffers from the same failure of decentralisation already discussed in the Ethererum technology stack, but this is compounded by the normalisation of intermediate art brokers \href{https://moxie.org/2022/01/07/web3-first-impressions.html}{continuing to custody} the NFTs even after sale. They are usually selling a pointer to their own servers. The market is nascent and evolving, but it's currently not delivering on it's core promise.\par
Proof of ownership is intuitively a pretty obvious application for the technology, but again it's hard to justify the expense when the benefits are so slim. \href{https://www.bullishlybred.com/}{Bulldogs on the blockchain} is a clear gimmick, and might even incentive poor behaviours as there are two products here which are not necessarily aligned. Much has been written over the years about \href{https://propy.com/browse/propy-nft/}{deeds to property} being passed through blockchains, cutting out the middle man, but in the event that a house deed NFT was hacked and stolen it's obviously not the case that the property would then pass to the hacker.

\section{Is any of this useful?}
Maybe, maybe not. This hype cycle isn't the best place or time to judge this.
\section{User stories / behaviours}
\label{behaviours}

\begin{itemize}
\item As a user I want to select a digital asset I find in a metaverse and then be offered an option to purchase the asset so I can look at it in my own spaces.
\item As a user I want to click on a digital asset I find in a metaverse and be given the opportunity to buy it as a rare digital representation so that only I and a few others are provably certified to own.
\item As a metaverse `land' holder (stakeholder) I would like to reward winners of competitions with real money or digital assets to foster and gamify the system
\item As a user I would like to interact with a virtual marketplace where I could swap and trade digital assets with other users so that I continue to feel engaged.
\item As a user I would like to create content so I can take it to metaverse locations and monetise it.
\end{itemize}

\section{Comparing the technologies}
Figure \ref{tab:compare} on the next page compares the technologies discussed so far in the context of the behaviours in section \ref{behaviours}. \par
Looking at these row by row it can be seen that:\par\
\begin{itemize}
\item Bitcoin layer 1 is too slow. Additionally the technical challenge of interacting with the network are considerable, and the fees are both high and likely to rise.
\item Ethereum ERC721 non fungible tokens might be a possibility for supporting limited edition digital assets. A web interface could be built on cloud servers to handle the shop and back end commitments to the Ethereum network. While fully committed transactions would take a few minutes to process the shop itself could inform the users immediately then email out a certificate of ownership. It would be possible to trade these items on other platforms, and conceivably within an app. Prizes (digital assets) could be issued to users for competition activity. The system would be expensive to use, the fees might vary wildly, and the ecological costs might be enormous. Batching transactions could overcome this, but that would rely on substantial take-up of the service, and management overhead.
\item Building and issuing and Ethereum based token, in an ICO process or similar seems a disproportionate amount of investment and would be expensive to use. 
\item A custom private blockchain offers no advantages over a database. It is a difficult thing to implement, but would have better performance and use cost to the users.
\item BTC pay server is an interesting open source project which allows companies to host a simple lightning / BTC web shop. It's not clear what the use case is here, but it is well engineered and scalable and should be born in mind when thinking about the other problems and systems.
\item Lightning network has open source connectivity into Unity, works well with QR codes, is fast, and has low transaction latency. The only issue is that it offers nothing but marketing hype over and above more traditional digital point of sale layers, and is more difficult to connect to money rails in the UK.
\item New layer 3 solutions like RGB, stacks, or Synonym might allow cheap and fast in
\end{itemize}

\section{User stories / behaviours}
\label{behaviours}

\begin{itemize}
\item As a user I want to.
\item As a user I want to.
\item As a user I want to.
\end{itemize}

\section{Comparing the technologies}
This section is well out of date and needs a tonne of work.\\
Figure \ref{tab:compare} on the next page compares the technologies discussed so far in the context of the behaviours in section \ref{behaviours}. \par
Looking at these row by row it can be seen that:\par\
\begin{itemize}
\item Bitcoin is too slow. Additionally the technical challenge of interacting with the network are considerable, and the fees are both high and rising.
\item Ethereum ERC721 non fungible tokens might be a possibility for supporting limited edition digital asset sales. A web interface could be built on cloud servers to handle the shop and back end commitments to the Ethereum network. While fully committed transactions would take a few minutes to process the shop itself could inform the users immediately then email out a certificate of ownership. It would be possible to trade these items on other platforms, and conceivable within the app. Prizes (digital assets) could be issued to users for competition activity. The system would be expensive to use, the fees might vary wildly, and the ecological costs might be enormous. Batching transactions could overcome this, but that would rely on substantial take-up of the service, and management overhead.
\item Building and issuing and Ethereum based token, in an ICO process or similar seems a disproportionate amount of investment and would be expensive to use. 
\item A custom private blockchain offers no advantages over a database. It is a difficult thing to implement, but would have better performance and use cost to the users.
\item BTC pay server is an interesting open source project which allows companies to host a simple lightning / BTC web shop. It's not clear what the use case is here, but it is well engineered and scalable and should be born in mind when thinking about the other problems and systems.
\item Lightning network has open source connectivity into Unity, works well with QR codes, is fast, and has low transaction latency. The only issue is that it offers nothing but marketing hype over and above more traditional digital point of sale layers, and is more difficult to connect to money rails in the UK.
\item Coloured coins would allow all of the desirable functionality, but it would be a substantial development challenge, using old open source code, and have high use fees. 
\item RGB MyCitadel stack seems highly appropriate and is open source but is in development and will not be available until later in the year.
\end{itemize}


\section{What are the options }
ETH, SOL, ADA, STX, RGB, Liquid?, RSK, so many others, this is gonna be exhausting to write about.
\input{NFTtable}
%\lipsum[50]
\section{Risks}
\subsection{Wage slavery / pyramid schemes}
Much of the current play to earn models are digital sweat shops. etc
\subsection{Energy concerns}
It is under discussed within the community that `minting' non fungible tokens, predominantly on Ethereum, is energy wasteful, on an already energy inefficient platform. As a random \href{https://carbon.fyi/?address=0x6Ec30Fd91A504Aad948839B985C7263888B2Ad68} {example this single collector of a few images} accounts for nearly 10 kilotonnes of carbon emissions. This is a subtly different problem to the security of the Bitdooin economic network in that these tokens are clealy non fungible, so cannot be reused again and again in order to reduce the unit environmental impact. 
\section{Why choose bitcoin again }
For the purposed of this book, we are going to once again explore Bitcoin as the platform of choice. To be clear it's not the popular, or intuitive choice. It's barely possible to mint NFTs on Bitcoin, and the community is deeply sceptical about it.\par
because it's the true opensource money option, the cost / barrier to entry is far lower than ethereum, and with new technology such as \href{https://diba.io}{NFT integration directly with DIBA on RGB this integrates}. 
%\lipsum[50]